Emittance preservation of an electron bunch in a loaded quasi-linear plasma wakefield

### 2017-07-28 Tex/Paper.tex
Daily commit
Added more text and figures

The baseline AWAKE drive beam current is insufficient to reach the non-linear regime and produce a plasma bubble. The
plasma electrons are depleted to around $65\%$ of nominal plasma density at the injection point of the electron beam.
This condition is replicated in out single bunch case as the peak density of the drive beam is $0.83\cdot n_{0}$. The
single bunch setup uses the baseline proton energy $W_{pb} = 400\unit{GeV}$, and retains the transverse size
$\sigma_{r} = k_{pe}$. The length of the bunch $\sigma_{z} = 40\unit{\mu m}$.

@@

The witness beam in our simulation differs from AWAKE baseline parameters on several key points. Initial beam energy is
set such that $\gamma_{eb} = \gamma_{pb} = 426.3$. This was done to eliminate the problem of initial de-phasing of the
witness beam. AWAKE baseline energy for the long witness beam for Run 1 is $W_{eb} = 10-20\unit{MeV}$
\cite{gschwendtner:2016}. Beam loading of a short witness beam is sensitive to initial position. AWAKE Run 2 will likely
require a higher initial energy, but a $W_{eb} = 30-50\unit{MeV}$ is sufficient to minimise de-phasing as the drift
$\Delta\xi \propto \gamma^{-2}$ \cite{berglyd_olsen:2015}. For this case we have eliminated it entirely in order to
reduce the number of variables in our parameter scans.

@@

\begin{figure}[hbt]
    \includegraphics[width=0.99\linewidth,trim={4mm 0mm 4mm 0mm},clip]{figures/beamLoading}
    \caption{\label{Fig:BeamLoading} Comparison between the unloaded longitudinal e-field (no witness beam) and the
        loaded e-field along the axis. The magnitude of the beam density along the axis is shown for reference. The
        beams travel towards the left. $\xi = z - tc$ is the position in the simulation box. The beam density is in
        units of initial plasma density $n_{0} = 7\nexp{-14}\unit{cm}^{-3}$. The bottom plot shows the plasma density
        for the loaded and unloaded case. The plasma depletion region behind the proton bunch has a minimum of $67.4\%${}
        of $n_{0}$. }
\end{figure}

\begin{figure}[hbt]
    \includegraphics[width=0.99\linewidth,trim={4mm 0mm 4mm 0mm},clip]{figures/plasmaDenTWake}
    \caption{\label{Fig:PlasmaDenTWake} The plasma density in grey with the proton beam (blue) and the electron beam
        (red) superimposed. The line plot indicates the transverse wakefield gradient $dW_{x}/dx$ where
        $W_{x} = E_{x} - v_{b} B_{y}$, evaluated along the beam axis.}
\end{figure}

@@

The single drive beam setup is designed to behave similarly to the self-modulated case. However, since the drive beam is
prevented from significant transverse evolution, we are presented with an idealised case where the electron witness beam
sees consistent fields. The field generated by the proton drive bunch is seen as the dotted lint in figure
\ref{Fig:BeamLoading}. With a proton beam density $n_{pb} \simeq n_{0}$, we are in the quasi-linear regime but near
non-linear \cite{rosenzweig:2010}. There is some depletion of plasma electrons, but not enough to form a bubble. The
dotted green line in the lower part of figure \ref{Fig:BeamLoading} shows that the on-axis plasma density has a
depletion of $67\%$.

@@

@@

@@

>>>>

The baseline AWAKE drive beam current is insufficient to reach the non-linear regime and produce a plasma bubble. The
plasma electrons are depleted to around $65\%$ of nominal plasma density at the injection point of the electron beam.
This condition is replicated in out single bunch case as the peak density of the drive beam is $0.83\cdot n_{0}$. The
single bunch setup uses the baseline proton energy $W_{pb} = 400\unit{GeV}$, and retains the transverse size
$\sigma_{r} = k_{pe} = 200\unit{\mu m}$. The length of the bunch $\sigma_{z} = 40\unit{\mu m}$.

@@

The witness beam in our simulation differs from AWAKE baseline parameters on several key points. Initial beam energy is
set such that $\gamma_{eb} = \gamma_{pb} = 426.3$. This was done to eliminate the problem of initial de-phasing of the
witness beam. AWAKE baseline energy for the long witness beam for Run 1 is $W_{eb} = 10-20\unit{MeV}$
\cite{gschwendtner:2016}. Beam loading of a short witness beam is sensitive to its position relative to the field
\cite{tzoufras:2009}. AWAKE Run 2 will likely require a higher initial energy, but a $W_{eb} = 30-50\unit{MeV}$ is
sufficient to minimise de-phasing as the drift $\Delta\xi \propto \gamma^{-2}$ \cite{berglyd_olsen:2015}. For this case
we have eliminated it entirely in order to reduce the number of variables in our parameter scans.

@@

\begin{figure}[hbt]
    \includegraphics[width=\linewidth,trim={2mm 0mm 2mm 0mm},clip]{figures/beamLoading}
    \caption{\label{Fig:BeamLoading} The top plot compares the unloaded longitudinal e-field (no witness beam, blue.
        dashed  line) and the loaded e-field (blue line) along the beam axis. The magnitude of the beam density along
        the axis is shown (red line) for reference. The bottom plot compares the plasma density along the beam axis for
        a drive beam with no witness beam (dashed green line), witness beam with no drive beam (dash-dotted green line),
        and both witness and drive beam present (continuous green line). $\xi = z - tc$ is the position in the
        simulation box and both beams travel towards the left. The beam and plasma densities arein units of initial
        plasma density $n_{0} = 7\nexp{-14}\unit{cm}^{-3}$.}
\end{figure}

\begin{figure}[hbt]
    \includegraphics[width=\linewidth,trim={2mm 0mm 2mm 0mm},clip]{figures/plasmaDenTWake}
    \caption{\label{Fig:PlasmaDenTWake} The plasma density in grey with the proton beam (blue) and the electron beam
        (red) superimposed. The line plot indicates the transverse wakefield gradient $dW_{x}/dx$ where
        $W_{x} = E_{x} - v_{b} B_{y}$, evaluated along the beam axis.}
\end{figure}

@@

The single drive beam setup is designed to behave similarly to the self-modulated case. However, since the drive beam is
prevented from significant transverse evolution, we are presented with an idealised case where the electron witness beam
sees consistent fields throughout the plasma stage. The field generated by the proton drive bunch is seen as the dotted
line in figure \ref{Fig:BeamLoading}. With a proton beam density $n_{pb} \simeq n_{0}$, we are in the quasi-linear
regime but near non-linear \cite{rosenzweig:2010}. There is some depletion of plasma electrons, but not enough to form a
bubble. The dashed green line in the lower part of figure \ref{Fig:BeamLoading} shows that the on-axis plasma density
has a depletion of $67\%$.

@@

The witness bunch generates its own wakefield, which in the accelerating phase cancels out the $E_{z}$ field generated
by the drive bunch. For a finite length bunch this causes the particles in the tail of the witness bunch to be
accelerated less than those at the front resulting in high energy spread in exchange for high energy transfer
\cite{van_der_meer:1985}. Since $E_{z} \propto \rho_{b}$, a beam profile $\rho_{b}(\xi)$ that perfectly flattens the
accelerating field within the witness beam, reduces energy spread to zero. The ideal shape is triangular or trapezoidal
with the peak density in the direction of the beam velocity \cite{katsouleas:1987, tzoufras:2009}.

@@

For Gaussian beams a reasonable flat field and consequently low energy spread can be achieved by controlling the charge
density. In our case, with a matched beam to the plasma density \ref{EQ:Matched}, we are limited on how much total
charge we can accelerate. The test case shown in figures \ref{Fig:BeamLoading} and \ref{Fig:PlasmaDenTWake} has a total
charge of $100\unit{pC}$ with a $\sigma_{r}=5.23\unit{\mu m}$ matching an initial normalised emittance
$\epsilon_{0} = 2\unit{\mu m}$.

\begin{figure}[hbt]
    \includegraphics[width=\linewidth,trim={2mm 0mm 2mm 0mm},clip]{figures/beamPhaseSpace}
    \caption{\label{Fig:BeamPS} Phase space charge distribution of a $100\unit{pC}$, $60\unit{\mu m}$ long witness beam
        after $4\unit{m}$ of plasma. Mean momentum is $1.67\unit{GeV/c}$ with an RMS spread of $87\unit{MeV/c}$
        ($5.2\%$) for the full beam.}
\end{figure}

As illustrated in figure \ref{Fig:BeamPS}, our base case of a $100\unit{pC}$, $60\unit{\mu m}$ long witness beam indeed
shows a difference in energy transfer to the tail of the beam. Also the front of the beam has a large tail in momentum
space. The positioning of the witness beam is chosen to put the bulk of the charge as close to the peak accelerating
field as possible, as can be seen in figure \ref{Fig:BeamLoading}.

Due to the high charge density resulting from a very narrow beam, the witness beam's own wakefield reaches the full
non-linear regime with $n_{eb} \approx 35\times n_{0}$. The plasma region reaches full depletion close to the peak of
the electron beam, which in turn generates strong focusing fields, with gradients of $20\unit{kT/m}$ near the beam axis
(see figure \ref{Fig:PlasmaDenTWake}).

@@

\section[\label{Ack}]{Acknowledgements}
% ******************************************************************************************************************** %

The simulations forming the basis for this study have been made possible using the open source version of QuickPIC
released in early 2017 and owned by UCLA.

These numerical simulations have been made possible through access to the Abel computing cluster in Oslo, Norway
maintained by UNINETT Sigma2 AS and financed by the Research Council of Norway, the University of Bergen, the University
of Oslo, the University of Tromsø and the Norwegian University of Science and Technology. Project code: nn9303k.

The authors would also like to acknowledge the OSIRIS Consortium for providing access to the OSIRIS framework. OSIRIS
was used extensively for simulations leading up to the work presented in this paper.
